### 2024-06-19 basic-statistics.qmd
20240619 - additions

- the sample size (*N*): the larger the *N*, the greater the power
- the [effect size](#sec-practicalSignificance): the larger the effect, the greater the power
    - for group comparisons:
        - larger between-group variance, and
        - smaller within-group variance (i.e., strong measurement precision)
- the alpha level (typically .05): the higher the alpha level, the greater the power; however, the higher we set the alpha level, the higher the likelihood of Type I errors (false positives)
- one- versus two-tailed tests: one-tailed tests have higher power than two-tailed tests
- within-subject versus between-subject comparisons: within-subject designs tend to have greater power than between-subject designs

>>>>

- the sample size (*N*): the larger the *N*, the greater the power
    - for group comparisons, the power depends on the sample size of each group
- the [effect size](#sec-practicalSignificance): the larger the effect, the greater the power
    - for group comparisons, larger effect sizes reflect:
        - larger between-group variance, and
        - smaller within-group variance (i.e., strong measurement precision, i.e., [reliability](#sec-reliability))
- the alpha level: the researcher specifies the alpha level (though it is typically set at .05); the higher the alpha level, the greater the power; however, the higher we set the alpha level, the higher the likelihood of Type I errors (false positives)
- one- versus two-tailed tests: one-tailed tests have higher power than two-tailed tests
- [within-subject](#sec-withinSubject) versus [between-subject](#sec-betweenSubject) comparisons: [within-subject designs](#sec-withinSubject) tend to have greater power than [between-subject designs](#sec-betweenSubject)
